Speculation in foreign exchange: noise or information?

### 2015-10-25 Knitr/ElsPaper.tex
Updated ElsPaper.tex Ready to be assessed by Andros.  Conclusion has been improved.

\begin{abstract}
If foreign exchange prices are driven by speculation based on noise rather than information, extreme levels of speculative sentiment and speculative activity should lead to price reversals. Informed speculation is sustainable as it is part of the price discovery process. A unique set of option prices measure speculative sentiment while regulatory positions determine the weight of speculative activity.  An event study shows that extremes of speculative sentiment or speculative activity do not increase the probability of price reversal, though variance-measured risk increases.  Price activity after the extreme is close to a random walk. This supports the view that speculation is informed and part of the process of price-discovery rather than uninformed noise.      
\end{abstract}
\begin{keyword}
JEL classification: G12, G14, G15; information, noise-trading, event studies.
\end{keyword}

@@

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to the movement of prices in the foreign exchange market\footnote{The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.  The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).}.  If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after extremes.

@@

\section{Speculation and foreign exchange returns}
If foreign exchange speculation is uninformed noise, extreme speculation, whether measured by sentiment or weight of activity, will coincide with deviations from fundamental value; if speculation is informed, speculation is part of the process of price discovery and extremes will provide no information and future prices will follow a random walk. An event study showed that speculation is associated with price movement: exchange rates move in the direction of speculative sentiment and activity.  However, once extreme, prices are as likely to continue as to reverse.

The results that are recorded are not a function of the design of the event study.  The event window and the quantile that is chosen to measure the extreme can be changed.  Alternative measures produced very similar results. Alternative quantiles from 1\% and 99\% to 20\% and 80\% have also been assessed.  Figures \ref{fig:ES1}, \ref{fig:ES2}, \ref{fig:ES3} and \ref{fig:ES4} have much wider windows and do not find any alternative to the view that extreme speculation is followed by a random walk.

The covariance of speculative sentiment and speculative positions with the return indicates that this activity influences prices.  It is not random.  These results show that foreign exchange speculation, whether measured using either sentiment or the weight of positions, is an important part of the price discovery process.  The random walk after the extreme is indicative of foreign exchange prices absorbing most of the information that is available.

It is encouraging that these two very different measures of speculative activity produce very similar results and that the results hold over different time frames: one week for the option data and one week for the regulatory position data.  It is less clear whether the results would be reproduced across markets for other financial assets.    These results do not rule out uninformed speculation.  The Friedman effect from the Noise-trader model depends on there being \emph{noise-trader risk} or some limits to arbitrage.  The effect is not evident in this study of foreign exchange.  However, this is a large and very liquid market, other asset classes may have alternative institutional features that many not mean that these results can be extrapolated without important institutional questons being asked.

The \emph{Tobin Tax} would put a charge against financial transactions, particularly in foreign exchange.  Part of the argument depends on speculation being uninformed as the reduced activity that is caused by the tax would, it is argued, limit volatility and allow more long-term fundamental evaluations to become more prominent.  However, if speculation is informed and part of the price discovery process, this tax may reduce liquidity and increase volatility.

>>>>

\begin{abstract}
If foreign exchange prices are driven by speculation based on noise rather than information, extreme levels of speculative sentiment and speculative activity should lead to price reversals. Informed speculation is sustainable as it is part of the price discovery process. A unique set of option prices measure speculative sentiment while regulatory positions determine the weight of speculative activity.  An event study shows that extremes of speculative sentiment or speculative activity do not increase the probability of price reversal, though variance-measured risk increases.  Price activity after the extreme is close to a random walk. This supports the view that speculation is informed and part of the process of price-discovery rather than uninformed noise.      
\end{abstract}
\begin{keyword}
JEL classification: G12, G14, G15; event studies, information, noise-trading, speculation.
\end{keyword}

@@

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to the movement of prices in the foreign exchange market\footnote{\label{FX}The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.  The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).}.  If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after extremes.

@@

\section{Speculation and foreign exchange returns}
If foreign exchange speculation is uninformed noise, extreme speculation, whether measured by sentiment or weight of activity, will coincide with deviations from fundamental value; if speculation is informed, speculation is part of the process of price discovery and extremes will provide no information about future prices. An event study showed that speculation is associated with price movement: exchange rates move in the direction of speculative sentiment and activity.  However, once extreme, prices are as likely to continue as to reverse. Speculative sentiment and speculative activity influence foreign exchange prices as indicated by the covariance of speculative measures with returns.  These results show that foreign exchange speculation is an important part of the price discovery process.  The random walk after the extreme is indicative of foreign exchange prices absorbing most of the information that is available.

The results are not a function of the design of the event study.  The event window and the quantiles that are chosen to measure the extreme can be changed.  Alternative measures produced very similar results. Alternative quantiles from 1\% and 99\% to 20\% and 80\% have been assessed; Figures \ref{fig:ES1}, \ref{fig:ES2}, \ref{fig:ES3} and \ref{fig:ES4} have much wider windows and do not find any alternative to the view that extreme speculation is followed by a random walk.

It is encouraging that two very different measures of speculative activity produce very similar results and that the results hold over different time frames: one week for the option data and one week for the regulatory position data.  It is less clear whether the results would be reproduced across markets for other financial assets. The results do not rule out uninformed speculation.  The Friedman effect from the Noise-trader model, whereby the informed speculators take advantage of the space created by price pressure from the uniformed, depends on there being \emph{noise-trader risk} or some limits to arbitrage.  The effect is not evident in this study.  However, as discussed in Footnote \ref{FX}, the foreign exchange market is a particularly large and liquid market, other asset classes may have alternative institutional features that mean a different relationship or balance between the informed and uninformed and therefore different results.

This is important because there is no unambiguous view of speculative activity and because speculation is an important issues in a number of policy debates.  Most notably, the \emph{Tobin Tax} would put a charge against financial transactions, particularly in foreign exchange.  Part of the argument depends on speculation being uninformed as the reduced activity that is caused by the tax would, it is argued, limit volatility and allow more long-term fundamental evaluations to become more prominent.  However, if speculation is informed and part of the price discovery process, this tax may have the opposite effect to what is envisaged.  These results suggest that it could reduce liquidity and increase volatility.
